### 2017-08-14 paper/response.tex
More work on Kevin's comments

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
\usepackage{natbib}
\usepackage{microtype}

\renewcommand{\quote}{\em}

\title{Response to referees}
\author{Brendon J. Brewer}
\date{}

\begin{document}
\maketitle

%\abstract{\noindent Abstract}

% Need this after the abstract
\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt}

\section*{Response to John Skilling's comments}

\section*{Response to the other reviewer's comments}

{\quote
This is a very interesting paper that pushes state-of-the-art in both theory and practice.  As a result, there is much to learn here and it will take further effort for me to fully absorb the ideas presented in this paper.  As a result, this review is being written based on what I understand of this paper to date.  Perhaps there are places in which the author can clarify the presentation.}

Thanks for the positive comments. I have tried to clarify the presentation.

{\quote
Estimation of entropy (and other information-theoretic quantities) and estimation of the uncertainties of the estimates is a notoriously difficult problem.  I personally, would have benefited greatly from having access, or knowledge of, such a methodology years ago.  The paper should definitely be published after some minor edits.}

That's why I started thinking about this problem...

{\quote
1. The author presents a methodology for computing entropies, and other information-theoretic quantities, using Skilling's nested sampling algorithm.  It should be noted that Skilling himself advocates the estimation of the "entropy", I believe that it is the KL-divergence, as the estimate of the uncertainty in the estimated evidence.  This should be mentioned, and possibly discussed within the context of this present paper.}

{\quote
2. With regard to the algorithm to compute the depth, it is not clear what '+ +' refers to in the step in which the latest estimate is appended to the results.  Also at that step, do the square brackets refer to an array (of one element)?  Or do they refer to some kind of rounding, which I doubt.}

Sorry about the confusion. I was trying to represent the operation of
appending one new element (a depth estimate)
to an existing list, which in Python is {\tt +}
and in Haskell is {\tt ++}. The square brackets were meant to represent
the wrapping of a value within a list (to make a list of length one),
since an append function is probably a function of two lists.
All of this was too obscure. I have replaced it with
\begin{align}
\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)
\end{align}
which is hopefully more understandable.

{\quote
3. In Section 2, the author writes "1. Points can be generated from the prior..."  Please clarify what is meant by "points".  Does the author mean samples?}

Yes. I have changed points to samples.

{\quote
The author writes:
"The former measures the degree to which one question answers another ex ante, and is a function of two questions. The latter measures the degree to which a statement answered a question ex post, and is a function of a question and a statement."
which reveals that the author has really have been thinking about questions!}

I'm glad you liked it!

{\quote
The author also writes:
"The mutual information is another way of describing the relevance of the data to the parameters. its definition, and relation to other quantities"
As far as I am aware, this is a novel way of thinking about mutual information!  Very nice!
In addition, I did not know that it could be written as the expected value of KL over datasets, which is fascinating.}

Thanks. In most information theory textbooks, these things are written in
terms of abstract `random variables' $X$ and $Y$. One of my subsidiary
goals in this paper was to discuss this stuff in terms and notation more
familiar to Bayesians.

{\quote
4. It would be useful to follow (13) with the calculation written in your notation rather than Skilling's.}

I decided to do this the other way around, first writing it in my notation and
then switching to Skilling's, which is more convenient when the computational
algorithm is being discussed.

{\quote
5. After (13): Why does the implied distribution of $L$-values tend to be heavy-tailed?}

Imagine doing simple Monte Carlo to get $Z$, by drawing from the prior and
collecting likelihood values, which you'll want to average. A histogram of
those likelihood values will look ridiculous, and you might have to wait
a million years until you get enough points such that your average becomes
accurate, because it will be dominated by huge likelihoods which are seen
rarely. Most NS people talk about this in terms of the parameter space, but
it also has an interpretation in terms of a `probability distribution for
$L$' which is implied by $\pi$.

{\quote
6. Just before (14) and in (15): I do not understand the notation regarding the implicit likelihood boundary involving \\mathbb\{1\}}

I've added a bit of explanation of the indicator function notation.

{\quote
7. After (26) what is meant by "NS repetitions"?}

It's the number of reference particles
whose depth was estimated using an NS run. I've added a sentence to this effect.

{\quote
8. After (51): I do not understand the statement "The variance of this mutual information estimate was reduced by using a common sequence of reference points for the marginal and joint entropy runs, an instance of the ‘common random numbers’ technique 1"}

{\quote
9. It wasn't clear to me how the distance function (25) was chosen and why.}

It is just the Euclidean distance between two data vectors, each of which
contains 100 co-ordinates. I added a sentence to the paper to clarify this.

{\quote
10. I had some questions about how the algorithm was conceived.  I had a difficult time envisioning how the blue distribution played the role of the prior, and why.  Perhaps some explanation is necessary here since NS is usually presented to as to assume that parameter changes were implemented to maintain a uniform prior.}

I know a few groups use that convention (transform the parameter space so
the prior becomes uniform), but I never have, and there's no reason to do
that in general, as long as it's possible to explore $\pi$ then everything
should be fine.

{\quote
11.  It was not entirely clear to me how the algorithm was used to estimate the entropy of a Gaussian.  How was the posterior (24) arrived at.  The prior is normal.  But what about the likelihood?  And then how was the distance measure (25) computed?}

I have added an extra paragraph with the intent of explaining this in more
detail.

{\quote
There are a number of typos:

1. End of Section 1: "I sometimes write general formulae are written ..."}
Fixed.

{\quote
2. Missing citation in first sentence of Section 6}
I removed that citation.

{\quote
3. After (47): question mark missing in case ii)}
Fixed.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}