### 2017-08-14 paper/response.tex
Finalish changes

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
\usepackage{natbib}
\usepackage{microtype}
\usepackage{amssymb}
\usepackage{color}

\definecolor{orange}{rgb}{0.7, 0.25, 0}
\definecolor{green}{rgb}{0, 0.5, 0}
\definecolor{darkblue}{rgb}{0, 0, 0.7}

\renewcommand{\quote}{\em \color{orange}}

\title{Response to referees}
\author{Brendon J. Brewer}
\date{}

\begin{document}
\maketitle

%\abstract{\noindent Abstract}

% Need this after the abstract
\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt}

\section*{Response to John Skilling's comments}

{\quote
This is an interesting paper, which addresses how to compute entropy values directly, without explicit
evaluation of the underlying probability distribution. There are a few places where I feel that presentation
could be improved, which I list below. Nevertheless, I should note two particular points.

1. Definition of entropy}

I agree with Skilling that the fundamental definition of entropy involves
a base measure $q(x)$, and that relative entropies or KL divergences are the
appropriate objects to be working with. However, since (as he points out)
the algorithm more naturally involves the averages of log probabilities,
I have decided to leave most of the discussion in terms of Shannon entropies.
However, I have added extra discussion to the introduction, about how using
the Shannon entropy is equivalent to a uniform $q$, and cautioning that
this should be kept in mind. I have also added some text after the definition
of the log-volume of a ball of radius $r$, which clarifies where the implicit
assumption of a uniform $q(x)$ comes into the procedure.

{\quote
2. Possible future work
There is an infelicity at the heart of the author’s approach. The little target balls of radius $L$ need to be
small in order to avoid undue smudging of the pointwise values being estimated. Yet, as L shrinks, the
estimates lose accuracy because of the square-root-N variability of nested-sampling volumes. So the user
will need to compromise, and it’s possible that future work may discover how to ameliorate this.}

This is an excellent point, and something I have worried about. I have added
some discussion of this in Section 4 (search for the paragraph containing
`trade-off'). I think the current algorithm will still work well when the
thing we want the entropy of is low-dimensional, even if it's some function of
an underlying high-dimensional quantity.

{\quote
I suggest that applying the author’s method directly to the Kullback-Leibler form (equations 11 and
12) may offer a better way forward.}

It was not obvious to me how this would work (except perhaps from using a
distance function defined in the coordinate system in which $q$ is uniform).
I will think about it, and present any future work in this way if it is
sufficiently neat.

{\quote 
Near the end of page 9, the author argues against trying that
because “the estimate for a given dataset may be biassed”. Possibly so, though I would argue that the
dominant uncertainty with nested sampling is statistical, with numerical biasses (of sensible quantities)
being considerably less. Anyway, it’s not entirely clear in these nonlinear problems whether bias matters
much. I have in mind the Bayesian evidence Z, where an unbiassed estimate would give a seriously
biassed estimate of the logarithm log Z (which is what matters for Bayes factors), and vice versa.
The advantage of aiming at KL is that the statistic can be computed “for free” parasitically upon
a nested sampling run. In fact, I myself use apparent convergence of KL as the most reliable general
termination criterion I have yet found. (Hint to author: it converges quicker if it’s evaluated as the sum
$KL =
X \delta L
Z
log
L
Z$
over likelihood L, rather than using the corresponding weighted sum over prior mass shells $\delta X$ that’s
better for getting Z and the posterior.)}

It is not the bias that I was particularly worried about. I have added a
comment about the bias potentially being negligible in many applications.
What I was more concerned about was being able to integrate over nuisance
parameters as well. I can't see a way to get the prior-to-posterior
KL divergence {\em of nuisance parameters only} using standard NS, but I can
get it using my approach.

{\quote
To the editors: I am not suggesting that there’s anything wrong with the author’s proposals. We
advance by trying things, and even if something does become superseded — which this may not — it
1still gives helpful perspective.}

I agree. In fact, this paper has gone through a few distinct stages where I
started seeing better ways of doing certain things. It's possible that it
will be superseded quickly, but I do think it's an advance. I'd like to get
people thinking about Nested Sampling in different scenarios, i.e., where
pi and L are 'any distribution' and 'any scalar function' rather than necessarily
being prior and likelihood. We might find applications to things other than
data analysis and statistical mechanics by doing that.

{\quote
Minor improvements
1. Wrong English in last sentence of 1.1 (page 2).}

Fixed.

{\quote
2. “were” not “where” at beginning of 2 (page 2) — use brain not spellchecker.}

Fixed.

{\quote
3. I found the notation a bit confusing. $H(\theta | d)$ looks like a function of the posterior $\theta$’s at specific
data d, whereas it actually means the average hHi for unknown d. I agree that it’s unnecessarily fussy
to write the formally correct H(P (· | ·)) for H as a functional of the posterior as a function of arbitrary
arguments which readers are supposed to remember are parameters and data. However, angle brackets,
$hH(\theta | d)i$ or $H(\theta | d)i d$ would have helped me by giving a useful clue. They might also let the author
avoid the idiosyncratic $H(\theta | d = d 0 )$ by writing $H(\theta | d 0 )$ without the angle brackets for H at the specific
d 0 .
There are various overloaded definitions of H according to type of argument (see equations 8, 9, 10).
Introducing an explicit common format
X X
H(argument) $\equiv$
p($\theta$, d) log p(argument)
would bring these together and incidentally help to stress the basic point that the paper is addressing
averages of logarithms.}

I thought about this for a while, and decided not to change the notation much.
The reason is that my notation is closer to what is used in introductory
information theory texts. One usually sees things such as $H(Y|X=x)$
and $H(Y|X)$, where the latter is the expected value of the former (averaged
over $X$). This is the standard definition of conditional entropy.
I think my writing makes it clear what I mean.
However, in my idiosyncratic $H(\theta | d = d_0)$ I decided to change the
subscript 0 to a subscript obs for observed.

{\quote
4. At the top of page 5, I find it a bit nicer to initialise nested sampling with X 0 = 1 rather than
X 1 = t 1 . That also helps with the evaluation...}

Unfortunately we have competing tastes again at this point.
I don't like to include $X_0$ because there's no zeroth iteration of Nested Sampling.

{\quote
5. I realise that L is the active constraint during iterations, but I wonder if readers might get confused
by the use of symbol L as radius (mnemonic length?) whereas the standard usage has been L = likelihood.}

Yes, $L$ was for length, and it was unintentional that I had $L$ for both
likelihood and length. I've changed this to $r$ for radius throughout.

{\quote
6. Equation 27 is messy in form and mistaken in typography. My memorable formula for
the volume of a hypersphere is}

Fixed, and I've replaced the volume formula with the given (easier to read and
remember) suggestion.

{\quote
7. Near the bottom of page 7 (Algorithm 1), I was confused by
b
h $\leftarrow$ b
h + + [k/N ]
Why double +? Why the space between them? Do square brackets mean integer rounding? If so, should
it be [k]/N ?}

Sorry about the confusion. I was trying to represent the operation of
appending one new element (a depth estimate)
to an existing list, which in Python is {\tt +}
and in Haskell (my current flavour of the month language) is {\tt ++}. The square brackets were meant to represent
the wrapping of a value within a list (to make a list of length one),
since an append function is probably a function of two lists.
All of this was too obscure. I have replaced it with
\begin{align}
\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)
\end{align}
which is hopefully more understandable.

{\quote
8. In the caption of Figure 1, “around the red point” or “at the red ball” might be better than “at
the red point”.}

Changed.

{\quote
9. At equation 29,
1/100
should be
1/100.01
to conform to the 4-figure accuracy elsewhere.}

That was actually a mistake (I forgot to factor in the non-uniform prior),
rather than overzealous rounding. Thanks for noticing it.

{\quote
10. Below equation (45), I think ``$\ln(2 \times 10 ^{-5} )$'' would better be explained as ``ln(2L) = -10.820''.}

Fixed. It now says $\ln (2r)$.

{\quote
11. In the last paragraph of the text (page 19), there is a correct statement about how “to increase
one’s confidence in the results, [without getting] a guarantee”. It might be worth pointing out that there
can never be such a guarantee for any method that relies on pointwise evaluations, because there will
always be some risk of a dominant spike being missed.}

I've added such a sentence.

{\quote
12. In the references, my search for ref. 20 gave me “Statistics and Computing, 2017, 27, 219–236”.}

Somehow my bibliography file had become corrupted. I have fixed this.

\section*{Response to the other reviewer's comments}

{\quote
This is a very interesting paper that pushes state-of-the-art in both theory and practice.  As a result, there is much to learn here and it will take further effort for me to fully absorb the ideas presented in this paper.  As a result, this review is being written based on what I understand of this paper to date.  Perhaps there are places in which the author can clarify the presentation.}

Thanks for the positive comments. I have tried to clarify the presentation.

{\quote
Estimation of entropy (and other information-theoretic quantities) and estimation of the uncertainties of the estimates is a notoriously difficult problem.  I personally, would have benefited greatly from having access, or knowledge of, such a methodology years ago.  The paper should definitely be published after some minor edits.}

That's why I started thinking about this problem...

{\quote
1. The author presents a methodology for computing entropies, and other information-theoretic quantities, using Skilling's nested sampling algorithm.  It should be noted that Skilling himself advocates the estimation of the "entropy", I believe that it is the KL-divergence, as the estimate of the uncertainty in the estimated evidence.  This should be mentioned, and possibly discussed within the context of this present paper.}

I have mentioned that NS computes the prior-to-posterior KL divergence, and
that this has both an interesting interpretation and is used as a termination
criterion. This was added to the first paragraph of Section 3.

{\quote
2. With regard to the algorithm to compute tSorry about the confusion. I was trying to represent the operation of
appending one new element (a depth estimate)
to an existing list, which in Python is {\tt +}
and in Haskell is {\tt ++}. The square brackets were meant to represent
the wrapping of a value within a list (to make a list of length one),
since an append function is probably a function of two lists.
All of this was too obscure. I have replaced it with
\begin{align}
\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)
\end{align}
which is hopefully more understandable.he depth, it is not clear what '+ +' refers to in the step in which the latest estimate is appended to the results.  Also at that step, do the square brackets refer to an array (of one element)?  Or do they refer to some kind of rounding, which I doubt.}

Sorry about the confusion. I was trying to represent the operation of
appending one new element (a depth estimate)
to an existing list, which in Python is {\tt +}
and in Haskell is {\tt ++}. The square brackets were meant to represent
the wrapping of a value within a list (to make a list of length one),
since an append function is probably a function of two lists.
All of this was too obscure. I have replaced it with
\begin{align}
\widehat{\mathbf{h}} \leftarrow \textnormal{append}\big(\widehat{\mathbf{h}}, k/N\big)
\end{align}
which is hopefully more understandable.

{\quote
3. In Section 2, the author writes "1. Points can be generated from the prior..."  Please clarify what is meant by "points".  Does the author mean samples?}

Yes. I have changed points to samples.

{\quote
The author writes:
"The former measures the degree to which one question answers another ex ante, and is a function of two questions. The latter measures the degree to which a statement answered a question ex post, and is a function of a question and a statement."
which reveals that the author has really have been thinking about questions!}

I'm glad you liked it!

{\quote
The author also writes:
"The mutual information is another way of describing the relevance of the data to the parameters. its definition, and relation to other quantities"
As far as I am aware, this is a novel way of thinking about mutual information!  Very nice!
In addition, I did not know that it could be written as the expected value of KL over datasets, which is fascinating.}

Thanks. In most information theory textbooks, these things are written in
terms of abstract `random variables' $X$ and $Y$. One of my subsidiary
goals in this paper was to discuss this stuff in terms and notation more
familiar to Bayesians.

{\quote
4. It would be useful to follow (13) with the calculation written in your notation rather than Skilling's.}

I decided to do this the other way around, first writing it in my notation and
then switching to Skilling's, which is more convenient when the computational
algorithm is being discussed.

{\quote
5. After (13): Why does the implied distribution of $L$-values tend to be heavy-tailed?}

Imagine doing simple Monte Carlo to get $Z$, by drawing from the prior and
collecting likelihood values, which you'll want to average. A histogram of
those likelihood values will look ridiculous, and you might have to wait
a million years until you get enough points such that your average becomes
accurate, because it will be dominated by huge likelihoods which are seen
rarely. Most NS people talk about this in terms of the parameter space, but
it also has an interpretation in terms of a `probability distribution for
$L$' which is implied by $\pi$.

{\quote
6. Just before (14) and in (15): I do not understand the notation regarding the implicit likelihood boundary involving \\mathbb\{1\}}

I've added a bit of explanation of the indicator function notation.

{\quote
7. After (26) what is meant by "NS repetitions"?}

It's the number of reference particles
whose depth was estimated using an NS run. I've added a sentence to this effect.

{\quote
8. After (51): I do not understand the statement "The variance of this mutual information estimate was reduced by using a common sequence of reference points for the marginal and joint entropy runs, an instance of the ‘common random numbers’ technique 1"}

I have tried to elaborate on this at the end of Section 7.

{\quote
9. It wasn't clear to me how the distance function (25) was chosen and why.}

It is just the Euclidean distance between two data vectors, each of which
contains 100 co-ordinates. I added a sentence to the paper to clarify this.

{\quote
10. I had some questions about how the algorithm was conceived.  I had a difficult time envisioning how the blue distribution played the role of the prior, and why.  Perhaps some explanation is necessary here since NS is usually presented to as to assume that parameter changes were implemented to maintain a uniform prior.}

I know a few groups use that convention (transform the parameter space so
the prior becomes uniform), but I never have, and there's no reason to do
that in general. As long as it's possible to explore $\pi$, then everything
will be fine.

{\quote
11.  It was not entirely clear to me how the algorithm was used to estimate the entropy of a Gaussian.  How was the posterior (24) arrived at.  The prior is normal.  But what about the likelihood?  And then how was the distance measure (25) computed?}

I have added an extra paragraph with the intent of explaining this in more
detail.

{\quote
There are a number of typos:

1. End of Section 1: "I sometimes write general formulae are written ..."}

Fixed.

{\quote
2. Missing citation in first sentence of Section 6}

I removed that citation.

{\quote
3. After (47): question mark missing in case ii)}

Fixed.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}